Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-27 fully-abstract-comp/technical-report.tex
Fix typos, add a comment.

We assume functions $\decPerm{}$ and $\encPerm{}$.
\lau{15-09-2017: TOOD, write more about these. Should be like in the local cap setting.}

@@

A program needs access to the following:
\begin{description}
\item[Code] The instructions of a program should of course be available so they can be executed.
\item[Program seals] Each call/return point of the program needs a unique seal in order to enforce well-bracketedness. This seal needs to be unique on the entire machine.
\item[Stack] A program stores the value of local variables on the stack. 
\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory seperate from the stack.
\item[Linking table] A progam may need to call other programs. Other programs are made available in a linking table.
\end{description}

@@

The \emph{code}, \emph{program seals}, and \emph{linking table} are available through a normal read-execute capability in the $\pc$ register. The program seals are found on the first address governed by this capability. The second address governe by the capability contains a normal read capability that governs the linking table and points to the first address. The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.

@@

The \emph{peristent data storage} is expected to be accisble through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register). There are two ways for a program to get this data capability: (1) the data part of the code/data part should be the sealed data capability. (2) During a call, the data capability should be stored on the stack, so it can be retrieved afterwards.
\begin{figure}
  \centering
  \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
  \caption{Linking and program layout.}
  \label{fig:trg-prog-link}
\end{figure}

\subsection{Compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]

@@

\clearpage
\section{Examples}
\subsection{Capability Opacity}
\label{subsec:capability-opacity}
This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions. The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.

@@

\section{Notes}
\subsection{Leuven stay conclusions}
\begin{enumerate}
\item Enter capabilities replaces by sealed code/data pairs \label{item:first-point}
  \begin{itemize}
  \item To allow us to forbid dynamic code generation
  \end{itemize}
\item Conditional full-abstraction
  \begin{itemize}
  \item No undefined behaviour by \emph{trusted code} in any context implies full-abstraction. (Blame idea: undefined as undef of current pc, so current pc can be checked for whether it was the trusted code).
  \item Avoids dynamic checks to protect trusted code against itself.
  \item Possible due to point \ref{item:first-point}.
  \end{itemize}
\item replace push/pull/call/ret/sload by symbolic return pointer (pair) and stack pointer.
  \begin{itemize}
  \item Allows backtranslation to be embedding into source language.
    \begin{itemize}
    \item Interpreter for backtranslation is not able to accurately replicate code length.
    \end{itemize}
  \end{itemize}
\item Worlds similar to CSF paper, but invariants on seals as well.
  \begin{itemize}
  \item Invariants says what can be sealed with a certain seals.
  \end{itemize}
\item Linker: resolves symbols
  \begin{itemize}
  \item Export refs (Question: do we allow other things than sealed code/data pairs to be exported?)
  \item Import refs
  \item Fresh seal requirement
  \end{itemize}
\item Components are memory segments and symbols.
\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, se example in subsection~\ref{subsec:capability-opacity}).
  \begin{itemize}
  \item Notice x86 allows instructions of up to infinite length, so this is not weird.
  \end{itemize}
\item Have opaque capabilities.
  \begin{itemize}
  \item Without opaque capabilities, stack consumption could be inferred through stack pointer index etc.
  \end{itemize}
\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a = a \text{ and } b = b$)
  \begin{itemize}
  \item Define one region that governs the stack
  \item Make the regions more general
  \item For now, we have not done this: instead we stick with the transition systems we have and have to regions governing the stack reflecting the CSF work.
  \end{itemize}
\end{enumerate}

>>>>

We assume functions $\decPerm{}$ and $\encPerm{}$.
\lau{15-09-2017: TODO, write more about these. Should be like in the local cap setting.}

@@

A program needs access to the following:
\begin{description}
\item[Code] The instructions of a program should of course be available so they can be executed.
\item[Program seals] Each call/return point of the program needs a unique seal in order to enforce well-bracketedness. This seal needs to be unique on the entire machine.
\item[Stack] A program stores the value of local variables on the stack. 
\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory seperate from the stack.
\item[Linking table] A program may need to call other programs. Other programs are made available in a linking table.
\end{description}

@@

The \emph{code}, \emph{program seals}, and \emph{linking table} are available through a normal read-execute capability in the $\pc$ register. The program seals are found on the first address governed by this capability. The second address governed by the capability contains a normal read capability that governs the linking table and points to the first address. The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.

@@

The \emph{peristent data storage} is expected to be accessible through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register). There are two ways for a program to get this data capability: (1) the data part of the code/data pair (for either a callback or a global function in the linking table) should be the sealed data capability. (2) During a call, the data capability should be stored on the stack, so it can be retrieved afterwards.
\begin{figure}
  \centering
  \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
  \caption{Linking and program layout.}
  \label{fig:trg-prog-link}
\end{figure}

\subsection{Compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]
\dominique{Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions? Is there another function that compiles a single instruction to a list of instructions?}

@@

\clearpage
\section{Examples}
\subsection{Capability Opacity}
\label{subsec:capability-opacity}

\dominique{Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions. The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.

@@

\section{Notes}
\subsection{Leuven stay conclusions}
\begin{enumerate}
\item Enter capabilities replaces by sealed code/data pairs \label{item:first-point}
  \begin{itemize}
  \item To allow us to forbid dynamic code generation
  \end{itemize}
\item Conditional full-abstraction
  \begin{itemize}
  \item No undefined behaviour by \emph{trusted code} in any context implies full-abstraction. (Blame idea: undefined as undef of current pc, so current pc can be checked for whether it was the trusted code).
  \item Avoids dynamic checks to protect trusted code against itself.
  \item Possible due to point \ref{item:first-point}.
  \end{itemize}
\item replace push/pull/call/ret/sload by symbolic return pointer (pair) and stack pointer.
  \begin{itemize}
  \item Allows backtranslation to be embedding into source language.
    \begin{itemize}
    \item Interpreter for backtranslation is not able to accurately replicate code length.
    \end{itemize}
  \end{itemize}
\item Worlds similar to CSF paper, but invariants on seals as well.
  \begin{itemize}
  \item Invariants says what can be sealed with a certain seals.
  \end{itemize}
\item Linker: resolves symbols
  \begin{itemize}
  \item Export refs (Question: do we allow other things than sealed code/data pairs to be exported?)
  \item Import refs
  \item Fresh seal requirement
  \end{itemize}
\item Components are memory segments and symbols.
\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, se example in subsection~\ref{subsec:capability-opacity}).
  \begin{itemize}
  \item Notice x86 allows instructions of up to infinite length, so this is not weird.
  \end{itemize}
\item Have opaque capabilities.
  \begin{itemize}
  \item Without opaque capabilities, stack consumption could be inferred through stack pointer index etc.
  \end{itemize}
\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a \rightarrow a \text{ and } b \rightarrow b$)
  \dominique{I don't remember what we meant with PORs, but perhaps quasi-reflexive partial orders.  Not sure what the point of this was again... }
  \begin{itemize}
  \item Define one region that governs the stack
  \item Make the regions more general
  \item For now, we have not done this: instead we stick with the transition systems we have and have two regions governing the stack reflecting the CSF work.
  \end{itemize}
\end{enumerate}
